---
title: "COVID CANADA"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
editor_options: 
  chunk_output_type: inline
---

> Marcelo Ponce, Amit Sandhel (2020). covid19.analytics: An R Package to Obtain, Analyze and Visualize Data from the Corona Virus Disease Pandemic. URL <https://arxiv.org/abs/2009.01091>

## Analysis of current COVID data and predictions

```{r include=FALSE}
rm(list = ls())
```

```{r initialisation_librairies, include=FALSE, results="hide"}
library(ggplot2)
library(forecast)
library(fpp2)
library(TTR)
library(dplyr)
library(RCurl)
library(tidyverse)
library(covid19.analytics)
library(prophet)
library(lubridate)
library(dygraphs)

```

```{r constants}
predictionDays <- 90
country <- "Canada"
population <- 37590000
noCache <- TRUE
limNbCases <- 1500
limNbDeaths <- 2000
limNbDeathsYear <- 150000

```


```{r load_data, include=FALSE, results="hide"}
if (noCache) {
  covid <- read_csv("https://covid.ourworldindata.org/data/owid-covid-data.csv")
  na <- filter(covid, continent == "North America")
  canada <- filter(na, location == "Canada")
  usa <- filter(na, location == "United States")
  ag <- covid19.data(case = "aggregated")
  tsc <- covid19.data(case = "ts-confirmed")
  tsd <- covid19.data(case = "ts-deaths")
  tsca <- covid19.data(case = "ts-ALL")
  download <- getURL("https://msss.gouv.qc.ca/professionnels/statistiques/documents/covid19/COVID19_Qc_HistoHospit.csv", ssl.verifypeer = FALSE)
  hosp_quebec <- read.csv(text = download)
  hosp_quebec <- hosp_quebec %>% rename("Date" = "ï..Date")
  hosp_quebec$Date <- as.Date(hosp_quebec$Date, "%Y-%m-%d")
  hosp_quebec <- hosp_quebec[order(hosp_quebec$Date), ]
  download <- getURL("https://cdn-contenu.quebec.ca/cdn-contenu/sante/documents/Problemes_de_sante/covid-19/csv/decesquotidien.csv", ssl.verifypeer = FALSE)
  deces_quebec <- read.csv(text = download, sep = ";")
  deces_quebec <- deces_quebec %>% rename("Date" = "ï..Date.de.dÃ.cÃ.s")
  deces_quebec <- deces_quebec %>% rename("Moyenne7jours" = "Moyenne.mobile.sur.7.jours..centrÃ.e.")
  deces_quebec$total_jour <- deces_quebec$CHSLD + deces_quebec$RPA + deces_quebec$Domicile.et.Inconnu + deces_quebec$RI.et.Autres

  download <- getURL("https://msss.gouv.qc.ca/professionnels/statistiques/documents/covid19/COVID19_Qc_RapportINSPQ_HistoVigie.csv", ssl.verifypeer = FALSE)
  confirmes_quebec <- read.csv(text = download)
  
  
}
```

```{r include=FALSE}
confirmes_quebec <- confirmes_quebec %>% rename(Date = ï..Date)
confirmes_quebec <- confirmes_quebec[-1, ]
```

```{r include=FALSE}
previsions <- function(dsDate, dsValeurs, titre, xlab, ylab, predDays) {
  dsInt <- dsDate
  y <- dsValeurs
  dfInt <- data.frame(dsInt, y)
  colnames(dfInt) <- c("ds", "y")

  m <- prophet(dfInt, yearly.seasonality = F, daily.seasonality = F)
  future <- make_future_dataframe(m, periods = predDays)
  forcast <- predict(m, future)
  plot(m, forcast, xlabel = xlab, ylabel = ylab) + ggtitle(paste(titre, predictionDays, " jours"))
}
```

```{r echo=FALSE}
print(paste("Data dated from: ", max(canada$date)))
```

Looking at if new cases can predict hospitalisation 13 days in future.

```{r 13joursTest}
# 13 jours Test
casCovidParJour <- confirmes_quebec
casCovidParJour <- casCovidParJour[order(casCovidParJour$Date), ]
casCovidParJour$Date <- as.Date(casCovidParJour$Date)
hospTreizeJours <- hosp_quebec

merg <- inner_join(casCovidParJour, hospTreizeJours, by = c("Date"))
merg <- merg[order(merg$Date), ]
# Remove start of pandemy irregular
merg <- merg[-1:-144, ]

newCases <- head(merg$Nb_Nvx_Cas, -13)
icu <- tail(merg$ACT_Si_RSS99, -13)
hosp <- tail(merg$ACT_Hsi_RSS99, -13)

corrNewCasesIcu <- data.frame(newCases, icu)
corrNewCasesHosp <- data.frame(newCases, hosp)

ratioIcu <- mean(newCases / icu)
ratioHosp <- mean(newCases / hosp)

# Mean ratio new cases / hosp + 13 days
# ratioHosp <- 1.97
# ratioIcu <- 10.6


m.icu <- lm(newCases ~ icu, data = corrNewCasesIcu)
coeffIcu <- m.icu$coefficients[2]
interceptIcu <- m.icu$coefficients[1]

m.hosp <- lm(newCases ~ hosp, data = corrNewCasesHosp)
coeffHosp <- m.hosp$coefficients[2]
interceptHosp <- m.hosp$coefficients[1]

merg$Date <- as.Date(merg$Date) + 13
last13Days <- tail(merg, 13)
last13Days$Pred_ACT_Hsi_RSS99 <- as.integer(interceptHosp + (last13Days$Nb_Nvx_Cas/coeffHosp))
last13Days$Pred_ACT_Si_RSS99 <- as.integer(interceptIcu + (last13Days$Nb_Nvx_Cas/coeffIcu))
last13Days <- last13Days %>% select(Date, Nb_Nvx_Cas, ACT_Hsi_RSS99, ACT_Si_RSS99, Pred_ACT_Hsi_RSS99, Pred_ACT_Si_RSS99)
last13Days$ACT_Hsi_RSS99 <- rev(last13Days$ACT_Hsi_RSS99)
last13Days$ACT_Si_RSS99 <- rev(last13Days$ACT_Si_RSS99)
```

Test de corrélation entre les nouveaux cas et le nombre de patients en soins intensifs 13 jours plus tard

```{r}
plot(newCases ~ icu, ylab = "Nouveaux cas", xlab = "Soins Internes 13 jours + tard")
```

```{r}
plot(rstudent(m.icu) ~ fitted(m.icu),
  ylab = "Résidus de Student",
  xlab = "Valeurs prédites",
  main = "Homogénéité des variances"
)
```

```{r}
qqnorm(rstudent(m.icu),
  ylab = "Quantiles observés",
  xlab = "Quantiles théoriques", main = "Normalité des résidus"
)
qqline(rstudent(m.icu))
```

```{r}
summary(m.hosp)
```

```{r}
summary(m.icu)
```



```{r}
cor.test(corrNewCasesIcu$newCases, corrNewCasesIcu$icu)
```

Test de corrélation entre les nouveaux cas et le nombre de patients hospitalisés 13 jours plus tard

```{r}
plot(newCases ~ hosp, ylab = "Nouveaux cas", xlab = "Hospitalisations 13 jours + tard")
```

Test de corrélation entre les nouveaux cas et le nombre de patients hospitalisés 13 jours plus tard

```{r}
cor.test(newCases, hosp)
```

```{r}

last13Days$ACT_Hsi_RSS99 <- rev(last13Days$ACT_Hsi_RSS99)
last13Days$ACT_Si_RSS99 <- rev(last13Days$ACT_Si_RSS99)

plot(last13Days$Date, last13Days$Pred_ACT_Hsi_RSS99, type = "l", col = "red", xlab = "Date", ylab = "Hospitalisations", main = "Prédiction Hospitalisations (13 Jours + Tard)", ylim = c(0, 1500))
lines(last13Days$Date, last13Days$ACT_Hsi_RSS99, col = "blue")
legend(c("bottomleft"), 0, legend = c("Pred + 13 jours", "- 13 jours"), col = c("red", "blue"), lty = 1:1, cex = 0.8)
text(last13Days$Date, last13Days$Pred_ACT_Hsi_RSS99 + 25, labels = as.character(last13Days$Pred_ACT_Hsi_RSS99))
text(median(last13Days$Date), 1200, paste("Moyenne +13 jours: ", as.integer(mean(last13Days$Pred_ACT_Hsi_RSS99))))
text(median(last13Days$Date), 1000, paste("Moyenne -13 jours: ", as.integer(mean(last13Days$ACT_Hsi_RSS99))))
```

```{r}
plot(last13Days$Date, last13Days$Pred_ACT_Si_RSS99, type = "l", col = "red", xlab = "Date", ylab = "Soins Intensifs", main = "Prédiction Soins Intensifs (13 Jours + Tard)", ylim = c(0, 300))
lines(last13Days$Date, last13Days$ACT_Si_RSS99, col = "blue")
legend(c("bottomleft"), 0, legend = c("Pred + 13 jours", "- 13 jours"), col = c("red", "blue"), lty = 1:1, cex = 0.8)
text(last13Days$Date, last13Days$Pred_ACT_Si_RSS99 + 25, labels = as.character(last13Days$Pred_ACT_Si_RSS99))
text(median(last13Days$Date), 250, paste("Moyenne +13 jours: ", as.integer(mean(last13Days$Pred_ACT_Si_RSS99))))
text(median(last13Days$Date), 200, paste("Moyenne -13 jours: ", as.integer(mean(last13Days$ACT_Si_RSS99))))
```

```{r}
plot(hosp_quebec$ACT_Si_RSS99, ylim = c(0, max(hosp_quebec$ACT_Si_RSS99)), xlim = c(0, nrow(hosp_quebec)), ylab = "Hospitalisations Soins Intensifs", xlab = "Jours", main = "Quebec: Actuel - Soins Intensifs")
```

```{r}
previsions(hosp_quebec$Date, hosp_quebec$ACT_Si_RSS99, "Québec - Prévision Soins Intensifs ", "Date", "Patients", 90)
```

```{r}
plot(hosp_quebec$ACT_Hsi_RSS99, ylim = c(0, max(hosp_quebec$ACT_Hsi_RSS99)), xlim = c(0, nrow(hosp_quebec)), ylab = "Hospitalisations", xlab = "Jours", main = "Quebec: Actuel - Hospitalisations")
```

```{r}

previsions(hosp_quebec$Date, hosp_quebec$ACT_Hsi_RSS99, "Québec - Prévision Hospitalisations ", "Date", "Patients", 90)
```

```{r}
plot(confirmes_quebec$Nb_Nvx_Cas, ylim = c(0, max(confirmes_quebec$Nb_Nvx_Cas)), xlim = c(0, nrow(confirmes_quebec)), ylab = "Cas", xlab = "Jours", main = "Quebec : Actuel - Cas")
```

```{r}
previsions(confirmes_quebec$Date, confirmes_quebec$Nb_Nvx_Cas, "Québec - Prévision - Cas ", "Date", "Cas", 90)
```

```{r}
plot(deces_quebec$total_jour, ylim = c(0, max(deces_quebec$total_jour)), xlim = c(0, nrow(deces_quebec)), ylab = "Décès", xlab = "Jours", main = "Quebec : Actuel - Décès")
```

```{r}
previsions(deces_quebec$Date, deces_quebec$total_jour, "Québec - Prévision - Décès ", "Date", "Décès", 90)
```

```{r}
plot(canada$total_deaths, ylab = "Nombre Morts", xlab = "Jour", xlim = c(0, nrow(canada)), ylim = c(0, 30000), main = paste(country, " - Total Morts"))
```

```{r}
plot(canada$total_cases, ylab = "Nombre Cas", xlab = "Jour", xlim = c(0, nrow(canada)), , ylim = c(0, max(canada$total_cases)), main = paste(country, " - Total Cas"))
```

```{r}
plot(canada$new_deaths, ylab = "Nombre Morts", xlab = "Jour", xlim = c(0, nrow(canada)), ylim = c(0, 400), main = paste(country, " - Morts / Jour"))
```

```{r}
plot(canada$new_cases, ylab = "Nombre Cas", xlab = "Jour", xlim = c(0, nrow(canada)), main = paste(country, " - Nouveaux Cas / Jour"))
```

```{r}
plot(canada$total_cases_per_million, ylab = "Number Deaths", xlab = "Day", xlim = c(0, nrow(canada)), , main = "Total Cases /1 Million Population", col = "green", ylim = c(0, limNbDeathsYear), cex.axis = 0.75)
points(usa$total_cases_per_million, col = "red")
legend(1, 40000, legend = c("US", "Canada"), col = c("red", "green"), lty = 1:1, cex = 0.8)
```

```{r}
plot(canada$total_deaths_per_million, ylab = "Number Deaths", xlab = "Day", xlim = c(0, nrow(canada)), main = "Total Deaths /1 Million Population", col = "green", ylim = c(0, limNbDeaths), cex.axis = 0.75)
points(usa$total_deaths_per_million, col = "red")
legend(1, 800, legend = c("US", "Canada"), col = c("red", "green"), lty = 1:1, cex = 0.8)
```

```{r}
plot(canada$new_cases_per_million, ylab = "Number Cases", xlab = "Day", xlim = c(0, nrow(canada)), main = "New Cases/1 Million Population per Day", col = "green", ylim = c(0, limNbCases), cex.axis = 0.75)
points(usa$new_cases_per_million, col = "red")
legend(1, 800, legend = c("US", "Canada"), col = c("red", "green"), lty = 1:1, cex = 0.8)
```

```{r}
plot(canada$new_deaths_per_million, ylab = "Number Cases", xlab = "Day", xlim = c(0, nrow(canada)), main = "New Deaths/1 Million Population per Day", col = "green", ylim = c(0, 15), cex.axis = 0.75)
points(usa$new_deaths_per_million, col = "red")
legend(1, 800, legend = c("US", "Canada"), col = c("red", "green"), lty = 1:1, cex = 0.8)
```

```{r}
tsCan <- ts(canada$new_deaths_per_million, start = 50, class = "ts")
plot(forecast(auto.arima(tsCan), h = 90), sub = "Forecast 90 jours")
```

```{r}
tsCan2 <- ts(canada$new_deaths, start = 50, class = "ts")
plot(forecast(auto.arima(tsCan2), h = 90), sub = "Forecast 90 jours", ylab = "Nb de morts par jour", xlab = "Jours")
```

```{r results="hide"}
tsc_us <- tsc %>% filter(Country.Region == country)
tsc_us <- data.frame(t(tsc_us))
tsc_us <- cbind(rownames(tsc_us), data.frame(tsc_us, row.names = NULL))
colnames(tsc_us) <- c("Date", "Confirmed")
tsc_us <- tsc_us[-c(1:4), ]
tsc_us$Date <- ymd(tsc_us$Date)
tsc_us$Confirmed <- as.numeric(tsc_us$Confirmed)
str(tsc_us)
```

```{r results="hide"}
tsd_us <- tsd %>% filter(Country.Region == country)
tsd_us <- data.frame(t(tsd_us))
tsd_us <- cbind(rownames(tsd_us), data.frame(tsd_us, row.names = NULL))
colnames(tsd_us) <- c("Date", "Confirmed")
tsd_us <- tsd_us[-c(1:4), ]
tsd_us$Date <- ymd(tsd_us$Date)
tsd_us$Confirmed <- as.numeric(tsd_us$Confirmed)
str(tsd_us)
```

```{r}
qplot(Date, Confirmed, data = tsc_us, main = paste("Covid19 - Number of confirmed Cases in ", country))
```

```{r}
qplot(Date, Confirmed, data = tsd_us, main = paste("Covid19 - Number of confirmed deaths in ", country))
```

```{r}
ds <- tsc_us$Date
y <- tsc_us$Confirmed
df <- data.frame(ds, y)
colnames(df) <- c("ds", "y")

ds2 <- tsd_us$Date
y2 <- tsd_us$Confirmed
df2 <- data.frame(ds2, y2)
colnames(df2) <- c("ds", "y")

tsd_us$delta <- c(NA, diff(tsd_us$Confirmed, lag = 1))
tsd_us$month <- format(as.Date(tsd_us$Date), "%m")
tsd_us$week <- week(as.Date(tsd_us$Date))
tsd_us$delta[is.na(tsd_us$delta)] <- 0
```

```{r}
dfChunked <- df[-(1:100), , drop = FALSE]
m <- prophet(dfChunked, yearly.seasonality = F, daily.seasonality = F)
future <- make_future_dataframe(m, periods = predictionDays)
forcast <- predict(m, future)
plot(m, forcast, xlabel = "Date", ylabel = "Cases") + ggtitle(paste(country, " - Predicted Covid Caseses ", predictionDays, " days"))
```

```{r}
dfChunked2 <- df2[-(1:100), , drop = FALSE]
m2 <- prophet(dfChunked2, yearly.seasonality = F, daily.seasonality = F)
future2 <- make_future_dataframe(m2, periods = predictionDays)
forcast2 <- predict(m2, future2)
plot(m2, forcast2, xlabel = "Date", ylabel = "Deaths") + ggtitle(paste(country, " - Predicted Covid Deaths ", predictionDays, " days"))
```

```{r}
dyplot.prophet(m, forcast,
  main = paste(country, " - Predicted Covid Cases ", predictionDays, " days - FB Prophet"),
  xlab = "Date", ylab = "Cases"
) %>% dyOptions(maxNumberWidth = 20)
```

```{r}
dyplot.prophet(m2, forcast2,
  main = paste(country, " - Predicted Covid Deaths ", predictionDays, " days - FB Prophet"),
  xlab = "Date", ylab = "Deaths"
) %>% dyOptions(maxNumberWidth = 20)
```

# FB Prophet

```{r}
prophet_plot_components(m, forcast)
```

## US Deaths

```{r}
prophet_plot_components(m2, forcast2)
```

```{r}
l <- nrow(dfChunked)
pred <- forcast$yhat[1:l]
actual <- m$history$y
plot(actual, pred, main = paste(country, " Cases - Prediction vs Actual"))
abline(lm(pred ~ actual), col = "red")
```

```{r}
pred2 <- forcast2$yhat[1:l]
actual2 <- m2$history$y
plot(actual2, pred2, main = paste(country, " Deaths - Prediction vs Actual"))
abline(lm(pred2 ~ actual2), col = "red")
```

```{r}
summary(lm(pred ~ actual))
```

```{r}
summary(lm(pred2 ~ actual2))
```

rmarkdown::render("COVID_CANADA.Rmd")
